\textbf{CSDS 491: Assignment 2}

### 2021-03-08 A2/article.tex
💄 Fix typo on email.
✨ Add answers for A2 Q1.

\author{Shaochen (Henry) ZHONG, ilc{sxz517@case.edu}}

@@

\section*{Q1}

\section*{Q2}

>>>>

\author{Shaochen (Henry) ZHONG, \ilc{sxz517@case.edu}}

@@

\section*{Q1. Conditional Independence}

\subsection*{1.1.}

\begin{align*}
    p(a,b) &= \sum\limits_c \sum\limits_d \sum\limits_e p(a, b, c, d, e) \\
    &= \sum\limits_c  p(c \mid a, b) p(a)p(b) \cdot \sum\limits_d p(d \mid c) \cdot \sum\limits_e p(e \mid c) \\
    &\text{Known that} \ \sum\limits_{x} p(x \mid pa(x)) = 1 \\
    &= p(a)p(b)
\end{align*}

Thus, $a \perp b \mid \emptyset$.

\subsection*{1.2.}

For $a \perp b \mid e$ we must have $p(a, b \mid e) = p(a \mid e) p(b \mid e)$. Which implies $p(a \mid b, e) = p(a \mid e)$ since:

\begin{align*}
    p(a \mid b) &= \frac{p(a, b)}{p(b)} \\
    p(a \mid b, e) &= \frac{p(a, b \mid e)}{p(b \mid e)} \\
    &= \frac{p(a \mid e) p(b \mid e)}{p(b \mid e)} \\
    &= p(a \mid e)
\end{align*}

Known that $a \perp b \mid \emptyset$, now to calculate $p(a \mid b, e)$ and $p(a \mid e)$ to see if they are equal:

\begin{align*}
    p(a \mid b,e) &= \frac{p(a,b,e)}{p(b,e)} = \frac{\sum_c p(a,b,c,e)}{\sum_a\sum_c p(a,b,c,e)}\\
    &= \frac{\sum_c p(e \mid c)p(a,b,c)}{\sum_a\sum_c p(e \mid c)p(a,b,c)}\\
    &= \frac{\sum_c p(e \mid c)p(a,b,c)}{\sum_a\sum_c p(e \mid c)p(c \mid a, b) p(a, b)}\\
    &= \frac{\sum_c p(e \mid c)p(a,b,c)}{\sum_a\sum_c p(e \mid c)p(c \mid a, b) p(a) p(b)}\\
    &= \frac{\sum_c p(e \mid c)p(a,b,c)}{\sum_c p(e \mid c)p(b, c)}
\end{align*}

\begin{align*}
    p(a \mid e) &= \frac{p(a,e)}{p(e)} = \frac{\sum_c\sum_b p(a,b,c,e)}{\sum_c p(e \mid c)p(c)}\\
    &= \frac{\sum_c p(e \mid c) \sum_b p(a,b,c)}{\sum_c p(e \mid c)p(c)}\\
    &= \frac{\sum_c p(e \mid c) p(a,c)}{\sum_c p(e \mid c)p(c)}
\end{align*}

Thus, it is possible that $p(a \mid b,e) \neq p(a \mid e)$ unless we may confirm that $p(b) = 1$.

\section*{Q2. Conditional Independence and Causality}

The conditional independence assumption this model made is $b \perp c \mid a$ as $p(b, c \mid a) = p(b \mid a) p(c \mid a)$; and joint distribution of this model is therefore $p(a, b, c) = p(b\mid  a)p(c\mid a)p(a)$.

However, the same joint distribution $p(a, b, c)$ can also be modeled differently:

\begin{align*}
    p(a, b, c) &= p(a \mid b, c)p(b \mid c)p(c) \\
    &= p(b \mid a, c)p(a \mid c)p(c) \\
    &= p(c \mid a, b)p(a \mid b)p(b) \\
    &= \dots
\end{align*}

Thus, the condition of the purposed graph does not necessarily hold.
